---
title: "Week 2"
subtitle: "Computational Sociology"
author: Christopher Barrie
format:
  revealjs:
    chalkboard: true
editor: visual
---

## Introduction

1.  Housekeeping
2.  Echo Chambers and Filter Bubbles

## Introduction: Echo Chambers

-   Who follows people they don't like on Twitter/Weibo/TikTok?

## Introduction: Echo Chambers

-   Who follows people they don't like on Twitter/Weibo/TikTok?

-   Who wants to give me their Twitter handle?

    -   How would you describe yourself politically?
    -   Would you believe me if I told you I already knew this?

## Introduction: Echo Chambers

```{r, echo = F}
library(academictwitteR)
library(ggplot2)
library(dplyr)

userfwing <- readRDS("data/cjbfwing.rds")
ids <- userfwing$id
```

```{r, eval = F, echo = T}
library(academictwitteR)
library(ggplot2)
library(dplyr)

get_user_id("cbarrie")
cjb_ID <- "95226101"
userfwing <- get_user_following(cjb_ID)
ids <- userfwing$id
```

## Introduction: Echo Chambers

```{r, echo = F}
library(tweetscores)
results <- readRDS("data/cjbtscore.rds")
plot(results)
```

```{r, eval = F}
devtools::install_github("pablobarbera/twitter_ideology/pkg/tweetscores")
library(tweetscores)

results <- estimateIdeology("cbarrie", ids)
plot(results)
```

## Introduction: Echo Chambers

Why am I showing you this?

-   Because *the reason* we can estimate ideology like this is because:

    -   We tend to follow people we agree with

    -   How come...?

## Dissonance is costly

Festinger (1957)

But is this the reason we follow people we agree with?

## Trust as alternative explanation

We might follow consonant opinion because we find it more trustworthy...

But why does this all matter for the internet

## Post-broadcast democracy

Prior and post-broadcast democracy; Sunstein and the Internet

But also we can be algorithmically manipulated

## Filter bubbles

Parisier and filter bubbles
